Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning

### 2018-05-16 nips_2018.tex
recognition model section

Imagine that you are asked to edit some text,
and told that you should change the text ``Nancy FreeHafer'' to ``Dr. Nancy''.
From this example,
you likely infer that ``Pushmeet Kohli'' should be changed to ``Dr. Pushmeet'',
drawing upon your prior knowledge of text,
like that words are separated by spaces and that
one commonly prepends titles like ``Dr.''
In this work,
we consider the problem of
building agents that solve few-shot learning tasks like these,
and also the problem of acquiring
the prior knowledge necessary to quickly
solve these tasks (Figure~\ref{firstPageFigure}).
We think of
solutions to these tasks as being best represented by programs,
and so our problem can be stated as follows:
how should an agent learn to write programs?
We take inspiration from two sources:
(1) Good software engineers compose libraries
of reusable subroutines
that are shared across
related programming tasks.
Returning to Figure~\ref{firstPageFigure},
a good string processing library should support appending strings and splitting on spaces --
exactly the prior knowledge needed to solve the task in Figure~\ref{firstPageFigure}.
(2) Skilled human programmers
can quickly recognize what kinds of programming idioms and library routines would be useful for solving
the task at hand, even if they cannot instantly work out the details.
\begin{figure}[h]
  \begin{minipage}[c]{0.68\textwidth}
\hspace{-0.5cm}      \begin{tabular}{r|c|}\cmidrule{2-2}
  \sc{Task}&
  \begin{tabular}{rcl}
  Nancy FreeHafer&$\longrightarrow$& Dr. Nancy\\
% Andrew Cencici&$\longrightarrow$&Dr. Andrew\\
 Pushmeet Kohli&$\longrightarrow$&???
  \end{tabular}\\\cmidrule{2-2}
  \sc{Program}&$f(\text{\code{s}})\,=\,$\code{(}$f_0$\code{ "Dr." (}$f_2$\code{ s " "))}\\\cmidrule{2-2}
  \begin{tabular}{c}
    \sc{Library}\\
    \sc{(DSL)}
    \end{tabular}&\begin{tabular}{l}
    $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b (lambda (x y)}\\
    \phantom{$f_0($\code{a}$,$\code{b}$) \,=\, $\code{(}}\code{(cons x y)))}\\
    \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
    $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (x a)}\\
    \phantom{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{  }}\code{(if (= c x) nil (cons x a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
    \end{tabular}\\\cmidrule{2-2}
  \end{tabular}
  \end{minipage}%
  \begin{minipage}[c]{0.31\textwidth}
      \caption{\textbf{Task}: Few-shot learning problem. Model solves tasks by writing \textbf{program}s, and jointly learns a \textbf{library} of reusable subroutines that are shared across multiple tasks, called a \textbf{Domain Specific Language (DSL)}. Program writing is guided by a neural network trained jointly with the library.}\label{firstPageFigure}
    \end{minipage}
\end{figure}
We combine these two ideas into
a new algorithm called \systemEnding,
which takes as input a collection of
programming \textbf{tasks},
and then jointly
solves three problems:
(1) Writing programs that solve the tasks;
(2) Composing a library of
domain-specific
subroutines -- which allow the agent to
more compactly write programs in the domain,
and (3) Training a neural network
to recognize
which library components are
useful for which kinds of tasks.
Together, the library and neural net
encode the domain specific knowledge needed to
quickly write programs.

@@

\textbf{Neural recognition model: tractably maxing $\lowerBound$ w.r.t. the
  frontiers.}  Here we train a neural network, $q$, to predict a
distribution over programs conditioned on a task. The objective of $q$
is to assign high probability to programs $p$ where
$\probability[x,p|\mathcal{D},\theta]$ is large, because including those programs
in the frontiers will most increase $\lowerBound$.  %% With $q$ in hand we can find programs for
%% the frontier $\mathcal{F}_x$ by searching for programs that maximize
%% $q(p|x)$.
%% Rather than directly predicting a distribution over $p$ conditioned on $x$,
%% the recognition model predicts a distribution $\theta^{(x)}$ over components of the DSL.
%% This taps into the intuition that programming is primarily a top-down activity:
%% as human programmers, we often first decide what kind of
%% programming constructs we might need to use,
%% and then we figure out how to assemble them in to the desired program.
%% This approach implements an amortized
%% inference scheme~\cite{le2016inference,ritchie2016deep} for the generative model in
%% Fig.~\ref{graphicalModel}.

@@

\noindent\textbf{Definition: $(\mathcal{D},\theta)$.}
A DSL $\mathcal{D}$ is a set of typed $\lambda$-calculus expressions.
%\\\noindent\textbf{Definition.}
A weight vector $\theta$ for a DSL $\mathcal{D}$ is a vector of $|\mathcal{D}| + 1$ real numbers:
one number for each DSL primitive $e\in \mathcal{D}$, written $\theta_e$ and controlling the probability of primitive $e$ occurring in a program,
and a weight controlling the probability of a variable occurring in a program, $\theta_{\text{var}}$.

@@

The purpose of the recognition model is to amortize the cost of searching for
programs.  It does this by learning to predict programs which are
probable under $(\mathcal{D},\theta)$ while also assigning high
likelihood to a task according to $\probability[x|p]$.  Concretely,
the recognition model $q$ is a neural network that predicts, for each
task $x\in X$, a weight vector $q(x) = \theta^{(x)}\in
\mathbb{R}^{|\mathcal{D}| + 1}$.  Together with the DSL, this defines
a distribution over programs, $\probability[p|\mathcal{D},\theta =
  q(x)]$.  We abbreviate this distribution as $q(p|x)$.  The crucial
aspect of this framing is that the neural network leverages the
structure of the learned DSL, so it is \emph{not} responsible for
generating programs wholesale.  We share this aspect with
DeepCoder~\cite{balog2016deepcoder} and~\cite{menon2013machine}.

We want a recognition model that closely approximates the true posteriors over programs. We formulate this as minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|q(p|x) \right) \right]$,
or equivalently maximizing $  \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]$,
%% \begin{equation*}
%%   \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]
%% \end{equation*}
where the expectation is taken over tasks. One could take this expectation
over the empirical distribution of tasks,
like how an autoencoder is trained~\cite{hinton2006reducing}; or, one could take this expectation over samples from the generative model, like how a Helmholtz machine is trained~\cite{dayan1995helmholtz}.
We found it useful to maximize both an autoencoder-style objective $\mathcal{L}_{\text{AE}}$ and a Helmholtz-style objective $\mathcal{L}_{\text{HM}}$, giving the  objective for a recognition model, $\mathcal{L}_{\text{RM}} = \mathcal{L}_\text{AE} + \mathcal{L}_\text{HM}$:
\begin{align}\nonumber
\mathcal{L}_{\text{HM}} = \expect_{(p,x)\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right]\quad
\mathcal{L}_{\text{AE}} = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
  \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\log q(p|x)\right]
\end{align}
The $\mathcal{L}_{\text{HM}}$ objective is essential for data efficiency:
all of our experiments train \system on only a few hundred tasks, which is too little for
a high-capacity neural network $q$.
Once we bootstrap a $(\mathcal{D},\theta)$,
we can draw unlimited samples from $(\mathcal{D},\theta)$
and train $q$ on those samples.

Evaluating $\mathcal{L}_{\text{HM}}$ involves sampling programs from
the current DSL, running them to get their outputs,
and then training $q$ to regress from the input/outputs to the program.
Since these programs map inputs to outputs,
we need to sample the inputs as well.
Our solution is to sample the inputs
from the empirical observed distribution of inputs in $X$.

>>>>

Imagine that you are asked to edit some text,
and told that you should change the text ``Nancy FreeHafer'' to ``Dr. Nancy''.
From this example,
you likely infer that ``Pushmeet Kohli'' should be changed to ``Dr. Pushmeet'',
drawing upon your prior knowledge of text,
like that words are separated by spaces and that
one commonly prepends titles like ``Dr.''
In this work,
we consider the problem of
building agents that solve few-shot learning tasks like these,
and also the problem of acquiring
the prior knowledge necessary to quickly
solve these tasks (Figure~\ref{firstPageFigure}).
We think of
solutions to these tasks as being well represented by programs,
and so our problem can be stated as follows:
how should an agent learn to write programs?
We take inspiration from two sources:
(1) Good software engineers compose libraries
of reusable subroutines
that are shared across
related programming tasks.
Returning to Figure~\ref{firstPageFigure},
a good string processing library should support appending strings and splitting on spaces --
exactly the prior knowledge needed to solve the task in Figure~\ref{firstPageFigure}.
(2) Skilled human programmers
can quickly recognize what kinds of programming idioms and library routines would be useful for solving
the task at hand, even if they cannot instantly work out the details.
\begin{figure}[h]
  \begin{minipage}[c]{0.68\textwidth}
\hspace{-0.5cm}      \begin{tabular}{r|c|}\cmidrule{2-2}
  \sc{Task}&
  \begin{tabular}{rcl}
  Nancy FreeHafer&$\longrightarrow$& Dr. Nancy\\
% Andrew Cencici&$\longrightarrow$&Dr. Andrew\\
 Pushmeet Kohli&$\longrightarrow$&???
  \end{tabular}\\\cmidrule{2-2}
  \sc{Program}&$f(\text{\code{s}})\,=\,$\code{(}$f_0$\code{ "Dr." (}$f_2$\code{ s " "))}\\\cmidrule{2-2}
  \begin{tabular}{c}
    \sc{Library}\\
    \sc{(DSL)}
    \end{tabular}&\begin{tabular}{l}
    $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b (lambda (x y)}\\
    \phantom{$f_0($\code{a}$,$\code{b}$) \,=\, $\code{(}}\code{(cons x y)))}\\
    \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
    $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (x a)}\\
    \phantom{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{  }}\code{(if (= c x) nil (cons x a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
    \end{tabular}\\\cmidrule{2-2}
  \end{tabular}
  \end{minipage}%
  \begin{minipage}[c]{0.31\textwidth}
      \caption{\textbf{Task}: Few-shot learning problem. Model solves tasks by writing \textbf{program}s, and jointly learns a \textbf{library} of reusable subroutines that are shared across multiple tasks, called a \textbf{Domain Specific Language (DSL)}. Program writing is guided by a neural network trained jointly with the library.}\label{firstPageFigure}
    \end{minipage}
\end{figure}
We combine these two ideas into
a new algorithm called \systemEnding,
which takes as input a collection of
programming \textbf{tasks},
and then jointly
solves three problems:
(1) Writing programs that solve the tasks;
(2) Composing a library of
domain-specific
subroutines -- which allow the agent to
more compactly write programs in the domain,
and (3) Training a neural network
to recognize
which library components are
useful for which kinds of tasks.
Together, the library and neural net
encode the domain specific knowledge needed to
quickly write programs.

@@

\textbf{Neural recognition model: tractably maxing $\lowerBound$ w.r.t. the
  frontiers.}  Here we train %% a neural network, $q$, to predict a
%% distribution over programs conditioned on a task. The objective of $q$
%% is
$q(p|x)$ to assign high probability to programs $p$ where
$\probability[x,p|\mathcal{D},\theta]$ is large, because including those programs
in the frontiers will most increase $\lowerBound$.  %% With $q$ in hand we can find programs for
%% the frontier $\mathcal{F}_x$ by searching for programs that maximize
%% $q(p|x)$.
%% Rather than directly predicting a distribution over $p$ conditioned on $x$,
%% the recognition model predicts a distribution $\theta^{(x)}$ over components of the DSL.
%% This taps into the intuition that programming is primarily a top-down activity:
%% as human programmers, we often first decide what kind of
%% programming constructs we might need to use,
%% and then we figure out how to assemble them in to the desired program.
%% This approach implements an amortized
%% inference scheme~\cite{le2016inference,ritchie2016deep} for the generative model in
%% Fig.~\ref{graphicalModel}.

@@

\noindent\textbf{Definition: $(\mathcal{D},\theta)$.}
A DSL $\mathcal{D}$ is a set of typed $\lambda$-calculus expressions.
%\\\noindent\textbf{Definition.}
A weight vector $\theta$ for a DSL $\mathcal{D}$ is a vector of $|\mathcal{D}| + 1$ real numbers:
one number for each DSL element $e\in \mathcal{D}$, written $\theta_e$ and controlling the probability of  $e$ occurring in a program,
and a weight controlling the probability of a variable occurring in a program, $\theta_{\text{var}}$.

@@

The purpose of training the recognition model is to amortize the cost of searching for
programs.  It does this by learning to predict, for each task,
programs with high likelihood according to 
$\probability[x|p]$ while also being
probable under the prior $(\mathcal{D},\theta)$.  Concretely,
the recognition model $q$ predicts, for each
task $x\in X$, a weight vector $q(x) = \theta^{(x)}\in
\mathbb{R}^{|\mathcal{D}| + 1}$.  Together with the DSL, this defines
a distribution over programs, $\probability[p|\mathcal{D},\theta =
  q(x)]$.  We abbreviate this distribution as $q(p|x)$.  The crucial
aspect of this framing is that the neural network leverages the
structure of the learned DSL, so it is \emph{not} responsible for
generating programs wholesale.  We share this aspect with
DeepCoder~\cite{balog2016deepcoder} and~\cite{menon2013machine}.

How should we get the data to train $q$?
This is nonobvious  because \system is only weakly supervised (i.e., learns only from tasks and not from (program, task) pairs).
One approach is to sample programs from the DSL,
run them to get their input/outputs,
and then train $q$ to predict the program from the input/outputs.
This approach is like how a Helmholtz machine
trains its recognition model during its ``sleep'' phase~\cite{dayan1995helmholtz}.
The advantage of ``Helmholtz machine'' training is that
we can draw unlimited samples from the DSL,
training on a large amount of data.
Another approach is
 self-supervised learning,
training $q$ on the (program, task)
pairs discovered by the algorithm so far.
The advantage of self-supervised learning
is that the training data is much higher quality,
because we are training on the actual tasks.
Due to these complementary advantages,
we train on both these sources of data.

Formally, $q$ should approximate the true posteriors over programs: minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|q(p|x) \right) \right]$,
 equivalently maximizing $  \expect[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) ]$,
%% \begin{equation*}
%%   \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]
%% \end{equation*}
 where the expectation is taken over tasks. Taking this expectation over the empirical distribution of tasks gives self-supervised training; taking it over samples from the generative model gives  Helmholtz-machine style training.
 The  objective for a recognition model ($\mathcal{L}_{\text{RM}}$) combines the Helmholtz machine ($\mathcal{L}_{\text{HM}}$) and self supervised ($\mathcal{L}_{\text{SS}}$) objectives, $\mathcal{L}_{\text{RM}} = \mathcal{L}_\text{SS} + \mathcal{L}_\text{HM}$:
\begin{align}\nonumber
\mathcal{L}_{\text{HM}} = \expect_{(p,x)\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right]\quad
\mathcal{L}_{\text{SS}} = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
  \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\log q(p|x)\right]
\end{align}
\begin{comment}
  The $\mathcal{L}_{\text{HM}}$ objective is essential for data efficiency:
all of our experiments train \system on only a few hundred tasks, which is too little for
a high-capacity neural network $q$.
Once we bootstrap a $(\mathcal{D},\theta)$,
we can draw unlimited samples from $(\mathcal{D},\theta)$
and train $q$ on those samples.
  \end{comment}
Evaluating $\mathcal{L}_{\text{HM}}$ involves sampling programs from
the current DSL, running them to get their outputs,
and then training $q$ to regress from the input/outputs to the program.
Since these programs map inputs to outputs,
we need to sample the inputs as well.
Our solution is to sample the inputs
from the empirical observed distribution of inputs in $X$.
